---
title: "Prediction of Exercise Quality"
output:
  html_document:
    toc: true
    theme: united
    fig_caption: yes
---

# Training and Validation Data

I first split the (training) data into two parts in the following manner:

```{r, echo=T}
library(caret)
data <- read.csv("./pml-training.csv", na.strings=c("NA", "", "#DIV/0!"))
set.seed(123)
inTrain <- createDataPartition(data$classe, p=0.6, list=F)
training  <- data[inTrain, ]
testing  <- data[-inTrain, ]
```

```{r, echo=F}
meaningful0 <- !(colnames(training) %in% c("X", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "num_window"))
meaningful1 <- !sapply(training,function(x) all(is.na(x)))
meaningful2 <- mapply( function(x,y) ifelse(!y, y, if(class(x) == "factor")  TRUE else var(x,na.rm=T) >abs(max(x, na.rm=T)) * 1e-10), training, meaningful0 & meaningful1)

knnImp <-  function(df, model=NULL){
    target.indices <- which(sapply(df, class) != "factor")
    targets <- df[,target.indices]
    non.targets <- df[,-target.indices]
    if(is.null(model)){
        res <- preProcess(targets,method="knnImpute")
    }else{
        res  <- model
    }
    df.part <- predict(res, targets)
    df <- cbind(non.targets,df.part)
    list(df=df, result=res)
}

conf.mat <- function(model, test.data){
    prediction <- predict(model,test.data)
    confusionMatrix(prediction,test.data$classe)
}
imputed <- knnImp(training[,meaningful0 & meaningful1 & meaningful2])
testing.reduced <- testing[,meaningful0 & meaningful1 & meaningful2]
test.1 <- knnImp(testing.reduced, imputed$result)
```

The data stored in variable _triaining_ is used to fit models and
the data stored in variable _testing_ is set aside for validation.

# Feature Selection

In selecting features, I removed several of the features and augmented them with the k-nearest-neighbor imputation. In the following subsection, I describe how I performed them.

Feature Removal
---------------

I removed the following features since all of their values are NA (e.g., empty strings or error message for "division by zero"):

- kurtosis_yaw_belt
- skewness_yaw_belt
- kurtosis_yaw_dumbbell
- skewness_yaw_dumbbell
- kurtosis_yaw_forearm
- skewness_yaw_forearm

Further, I also removed the following features since they are not useful or inappropriate.
For instance, given a user_name, raw_timestamp_part_1 is a perfect predictor for training and validation data set because a user's data belonging a class have neighboring timestamps.
The following figure depicts this fact and confirms that this feature should not be taken into modeling.

```{r, qplot, echo=F, fig.cap="Figure: Plot of raw_timestamp_part_1 vs. classe."}
qplot(classe, raw_timestamp_part_1, data=subset(training,user_name=="charles"), colour=classe)+theme_bw()
```

Feature Imputation
------------------

The authors of the paper[@mainPaper] reported in the section of feature selection: _"17 features were selected:
in the belt, were selected the mean and variance of the roll,
maximum, range and variance of the accelerometer vector,
variance of the gyro and variance of the magnetometer."_

To utilize this information, I decieded to impute the average and variance of features based on existing values. After extracting columns that are not a factor, I applied k-nearest-neighbor imputation using caret package with default parameters.


# Result

After preprocessing, I fit a random forest model to the data.
The confusion matrix of the resultant model is shown below:

```{r, echo=F}
load(file="mymodel")
model1 <- models[[1]]
conf.mat(model1,test.1$df)
```

# References

---
nocite: |
  @dataSet
---

---
references:
  - id: dataSet
    title: Weight Lifting Exercises Dataset
    URL: http://groupware.les.inf.puc-rio.br/har
  - id: mainPaper
    title: Qualitative Activity Recognition of Weight Lifting Exercises
    author:
      - given: Eduardo
        family: Velloso
      - given: Andreas
        family: Bulling
      - given: Hans
        family: Gellersen
      - given: Wallace
        family: Ugulino
      - given: Hugo
        family: Fuks
    URL: http://groupware.les.inf.puc-rio.br/public/papers/2013.Velloso.QAR-WLE.pdf
    issued:
      year: 2013
---
